---
title: "R markdown exercise"
author: "Joshua"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




1) Load bacteria.csv into a dataframe


```{r}
getwd()
setwd()
```



```{r}
getwd()
bacteria_data <- read.csv("bacteria.csv", header = T, sep = ",")

```
Check the internal structure of the dataframe

```{r}
head(bacteria_data)
colnames(bacteria_data)
summary(bacteria_data)
str(bacteria_data)

#Save the dataframe in another name.
bacteria_data1 <- bacteria_data
```


2) Get the position of the first occurence of "Francisella tularensis" in the dataframe.


```{r, message=FALSE}
#First check the column that has "Francisella tularensis"
#"Francisella tularensis" %in% bacteria_data1$genome

# Using match to get the first occurence of "Francisella tularensis"

match("Francisella tularensis", bacteria_data1$genome)

#to check all locations that have "Francisella tularensis"
which(bacteria_data1$genome %in% "Francisella tularensis")
```

3) Which different values are present in the column "isolation_country"?

```{r}
# to get a vector of unique values that are present in the column "isolation_country"
unique(bacteria_data1$isolation_country)

# to get a number of unique values that are present in the column "isolation_country"
table(bacteria_data1$isolation_country)

```


4) Draw a scatterplot to visualize the relationship between the columns "GC" and "contigs".
Save the plot as a pdf.

```{r}
#Scatter plot with color and proper label
plot(bacteria_data1$GC,bacteria_data1$contigs, col= as.numeric(as.factor(bacteria_data1$contigs)), 
     main = "scatterplot of GC and Contigs", xlab = "GC", ylab = "Contigs", cex = 2,pch=16)


#To save the plot as pdf
pdf("Bacteria_GC_Contigs.pdf")

plot(bacteria_data1$GC,bacteria_data1$contigs, col= as.numeric(as.factor(bacteria_data1$contigs)), 
     main = "scatterplot of GC and Contigs", xlab = "GC", ylab = "Contigs", cex = 2,pch=16)

dev.off()
```


5) Create a subset, removing cases with less than 1797796 bp ("bp" column).
How long is the new dataframe?

```{r}
# To get the range of values of a "bp" column
range(bacteria_data1$bp)

#To creat the subset
Bacteria_data_bp <- subset(bacteria_data1, bacteria_data1$bp < 1797796)

# To get the lenght of the new data frame.
length(Bacteria_data_bp)# lenght and ncol finds the number of colums
ncol(Bacteria_data_bp)
nrow(Bacteria_data_bp) #To get the number of rows
```

6) In the dataframe new data frame replace the occurences of "Medium" in the "quality" column with the new value "Low"

```{r}
# First inspect the "quality" column
table(Bacteria_data_bp$quality)


Bacteria_data_bp$quality <- gsub("Medium","Low", Bacteria_data_bp$quality)

#Inspect the new quality column
table(Bacteria_data_bp$quality)
```

7) Create two subset from the new dataframe: "HIGH" with all the cases with a "High" value in the "quality" column, and "LOW" with all the cases with a "Low" value in the "quality" column


```{r}
# Subset with "High"
High <- subset(Bacteria_data_bp, Bacteria_data_bp$quality == "High")

#Subset with "Low"
Low <- subset(Bacteria_data_bp, Bacteria_data_bp$quality == "Low")


```



8) Get the maximum and minumum value of "GC" in the dataframe "Low"

```{r}
#To get the range
range(Low$GC)

#To get min and max seperately
min(Low$GC)
max(Low$GC)
```

9.1) What kind of variables are the columns "quality" and "bp"?
Quantitative or qualitative?

```{r}
typeof(bacteria_data1$quality)
typeof(bacteria_data1$bp)

#OR

class(bacteria_data1$quality) #qualitative
class(bacteria_data1$bp) #quantitative
```


9.2) Test if there is a significant difference between the "bp" of "HIGH" and "LOW".

```{r}
shapiro.test(High$bp)
shapiro.test(Low$bp)

# Since the p-value of both data is less than 0.05, we assume that the distribution is not normal. Therefore, we use wilcox test

wilcox.test(High$bp, Low$bp)
```

9.3) Compare the p-value calculated with the p-value threshold.

## The p-value I got is higher than 0.05.


9.4) Comment on the results.

## Since the p-value calculated is higher than the p-value threshold, we assume that there is no significantly different between "bp" of "HIGH" and "LOW".
